---
title: "Final_R_reference"
author: "Nicole Yunes Perez"
date: "2022-12-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

```{r}
library(palmerpenguins)
install.packages(here)
require(here)

ginkgo<- read.csv(here("Eco602", "environmental_data", "data", "ginkgo_data_2021.csv"))

```
In this section, I have loaded my built in data (palmerpenguins), installed the "here" package, require said package and then used that function and read.csv to load it into a dataframe called gingko. with the here function you can specify in which folders the document lives

## Data Structures

### The C function
The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

All of the elements must be of the same type.
I can’t combine character and numeric types in the same call to c()
Here’s two examples using numeric and character data types:
```{r}
num_vec  = c(1, 4, 8, 9, 13)

char_vec = c("a", "fish", "data is cool")
```
I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
num_vec
```
[1]  1  4  8  9 13
```{r}
print(char_vec)
```
[1] "a"            "fish"         "data is cool"

###The Length Function
The function length() tells us the length of a vector. For example, when used with our previous vector:

```{r}
length(num_vec)

```
Here the function length() tells us that there are 5 numbers in this vector.

### The Matrix Function

This function allows us to build a matrix, with any specified number of rows and columns. You must specify which data you will use for this matrix.

```{r}
number_matrix<- matrix(num_vec, 1, 5)
char_matrix<- matrix(char_vec, 1, 3)
```

So in this example, I've created a matrix with our previous number vector and put those numbers into a matrix with 1 row and 5 columns. The number of rows and columns must be a multiple of the data length, in this case that is 5.

### The data.frame Function
```{r}
num_df<- data.frame(number_matrix, char_matrix)
```

This function allows us to create a dataframe and congregate multiple matrices togethe. This is usually the level which we perform data analysis at.

### Ginkgo nrow
```{r}
nrow(ginkgo)
```
This function is fairly simple, and it just tells us the number of rows in any object already defined. In this case it is 756 rows.

### Ginkgo ncol
```{r}
ncol(ginkgo)
```
This function tells us the number of columns in any specified object. In this case, there are 6 rows.

### Ginkgo dim
```{r}
dim(ginkgo)
```
The dim() function tells us both the number of rows and the number of columns. It first lists the number of rows and then the number of columns.

## Subsetting
### $ for subsetting
We can use the '$' symbol to subset a part of our dataframe using the name of the row or column. 
```{r}
seeds<- ginkgo$seeds_present
```
In this example we've subsetted only the column which tells us if there are any seeds present.
### [] for subsetting
We can also use brackets to specify a specific row/column or combination of those.
```{r}
first.row.ginkgo<- ginkgo[1,]
row.2.col.3.ginkgo<- ginkgo[2,3]
third.col.ginkgo<- ginkgo[,3]
```
If you would only like to select an entire row or column, you must use an empty space and comma to represent no row/column was selected.

##Numerical Data Exploration

###Summary Function
If you would like to see a quick summary of your data, the summary function gives you a quick snapshot of some simple data exploration of your dataframe.
```{r}
summary(ginkgo)
```
This function will give you a range, mean, mode, and median of all the matrices in your dataframe.

###Mean Function
This function is pretty self explanatory. It tells you the mean/average of your specified dataset. You must specify which column you would like a mean of if there are multiple columns in your dataset.
```{r}
mean(ginkgo$max_width)
```
In this example, I've specified that I'd like to know the mean of the max_width for my ginkgo dataset and R tells us the mean is 74.34.
You can also find this information from the summary function.
### Standard Deviation Function
This function tells us what the standard deviation of our data looks like. Again, you must specify which column you would like to look at.
```{r}
sd(ginkgo$max_width)
```
In this example, R tells us the standard deviation of our max_width is 16.42.
